Corporate Security Solutions for BYOD:\\ A Novel User-Centric and Self-Adaptive System

### 2015-07-09 Response to reviewers - r2.tex

\end{verbatim}

We are very grateful for the opportunity of improving the paper in this revised version.

Following the Editor's and some of the reviewers' requests, we have updated the state of the art, revising and commenting in the text (Section 2) several works with common topics with this manuscript. The new references can be consulted b the end of this letter.

In addition to this, we have addressed all the requests and suggestions by the reviewers. The responses are included below in this document.

@@

% -----------------------------------------------------------------------------

\section{Comments by Reviewer \#3}

\begin{verbatim}
This paper describes MUSES, an adaptable framework for enforcing 
security policies on mobile devices. This tool relies on Machine 
Learning and CI techniques to predict policy violations and is 
self-adaptive on user behaviors.
The idea seems interesting although several technical concerns 
arise from the reading of this work:
1.      The tools aims at enforcing security policies, however 
the security response occurs only AFTER a particular pattern is 
recognized to be potentially malicious. From my understandings, 
there is no a-priori enforcement except for the prediction based 
on event logs. This could mean that a security violation could 
in principle happen and only later recognized. This is not so 
suitable for a corporate environment with strict security 
requirements. This statement should be discussed and preferably 
discredited with experimental results.
\end{verbatim}

MUSES starts from a defined set of security rules that the CSO must
define. Ideally, these rules should deal, in principle, with all the
potential security incidents that the company aims to detect. MUSES
would be able to define new rules, by means of refinement, which could
be labelled, i.e. assigned a decision, after the corresponding events
have happened. However, it can also infer or create new rules, which
the CSO must approve, using computational intelligence
techniques. These rules could deal with unexpected situations, which
have not happened previously .

@@

\textit{However, MUSES presents a limitation regarding the enhancement of rules, since, in principle, it cannot predict (generate rules for dealing with) unexpected or unknown events which could lead to a security incident. The philosophy is that the initial set of rules, defined by the CSO, should be very restrictive regarding possible unexpected users' behaviours or events, in order to avoid as much security incidents as possible. As the system works, MUSES would be able to define new rules (through refinement) which could get an associated decision after the corresponding events have happened. Thus, this will lead to get eventually an optimal set of security rules.}

\textit{In addition, MUSES can also infer or create new rules using computational intelligence techniques. These rules could deal with unexpected situations (not previously happened), but must be previously approved by the CSO. Of course, everything is constrained by the available set of sensors which, in turn, define the possible information that MUSES will analyze and use in the refinement and inference processes.}

\begin{verbatim}
2.      A strong limitation is that an effective enforcement is 
only possible with MUSES Aware Apps. It is unclear the effectiveness 
of the approach with commercial applications. In page 14 "MUSES 
could just detect very specific events and also do very limited 
actions". With those assumptions, it seems unlikely that a company 
would adopt MUSES since it is not able to offer an adequate level of 
security. I think that this is very far from being acceptable for a 
BYOD solution. Also in this respect, experimental results are needed.
\end{verbatim}

\textcolor{blue}{TODO - Justificar su valÌa. Las empresas querr·n MUSES y querr·n que sus empleados utilicen MUSES-Aware Apps porque les garantizar·n un control sobre lo que se hace en la empresa y los posibles incidentes de seguridad.}

With respect to the experimental results, a new section (4.4, in page XX) has been included describing some trials conducted in an actual company, using a prototype of MUSES on Android devices. The results show the effectiveness of this system, which has roughly reduced from 23000 security violations in the first stages to 0 in a few weeks.

% Antonio - completar algo m·s, si alguien quiere

@@

\begin{verbatim}
3.      MUSES Sensors are essentially user-space applications that 
collects logs on specific activities. This aspect really concerns 
me since a compromised device or compromised sensor apps could vanish 
all MUSES effort. Authors should deeply discuss this aspect.
\end{verbatim}

To clarify this point, a more detailed definition and description of the sensors has been included in Section 4.1 (page XX). In addition, a discussion on how the case the reviewer mentions could affect the system has been added just below the list of sensors, as:

@@

\textit{It is important to note that MUSES does not work as an antivirus itself, but has a sensor which checks if the device has already an antivirus, and raises a security violation alert if not. It might happen then that the user ignores this advice and the device becomes compromised. As it would be explained in the next section, the sensors are constantly sending their monitored values, and therefore patterns are created, stored, and processed on the server. In case the device gets infected with rootkit-type malware \cite{bickford2010rootkits}, a sudden change in the values of the sensors will mean a significant change in the patterns. Then, this change in the patterns will be noticed by the server, which will immediately warn the user, and decrease the device trust value.}

\begin{verbatim}
4.      Paper structure is really confusing. Many essential concepts 
are introduced and explained too late in the paper forcing the reader 
to go back and forth to understand the architecture (e.g. MUSES aware 
apps and Data Mining techniques). I suggest authors to deeply revise 
the presentation of their work.
\end{verbatim}

Following the reviewer's suggestion, the paper structure has been redefined, with an introduction that presents the work, a section that introduces preliminary concepts followed by the state of the art in corporate mobile security and, eventually, the description of the framework that is the main point of the paper. Changes have been too extensive to include here, but great
care has been taken to talk about concepts only when defined
previously. For instance, the concept of MUSES-Aware apps is now
introduced in the context of the API in Section 3.5 and only used
later on, from Section 3.6.4, in the context of other tool. The
concept of data mining is presented along with machine learning in section 4.3.1. 
We have also added some tables (in Section 5) as an additional summary of information, for the sake of clarity.

% Antonio - Mejorar un poco esta justificaciÛn, si hay tiempo

@@

\begin{verbatim}
5.      A discussion on research activities on BYOD is completely 
missing. This is unacceptable for a solution that aims at solving 
BYOD security issues. Just to cite a few:
a.      Gessner, Dennis, et al. "Towards a User-Friendly 
Security-Enhancing BYOD Solution." NEC Technical Journal 
7.3 (2013): 113.
b.      Samaras, Vasileios. A BYOD Enterprise Security Architecture 
for accessing SaaS cloud services. Diss. TU Delft, Delft 
University of Technology, 2013.
c.      Armando, Alessandro, et al. "Enabling BYOD through secure 
meta-market." Proceedings of the 2014 ACM conference on Security 
and privacy in wireless and mobile networks. ACM, 2014.
\end{verbatim}

The reviewer is right. Although several works have been cited in other sections of the work, a proper discussion on research on BYOD was necessary. Therefore, we have analysed the papers suggested by the reviewer, along with several new relevant ones in Section 2 (page XX). These new references have been also cited in other sections of the work in order to justify MUSES features; for example, the set of best practices for BYOD security proposed by H. Romer.

These works have been also considered to propose the new taxonomy included in the paper (beginning of Section 3, page XX). These references are \cite{Garba15organisational,Scarfo12survey,Gessner13userfriendly,Armando14metamarket,Haejung12Door,Samaras13SaaS,Miller12Privacy,Romer14BestPractices}, and can be found at the end of this document.

@@

\begin{verbatim}
6.      The experimental section is completely missing. Is MUSES 
able to identify security violations? Which are the computational 
cost for both the company (server) and users (mobile devices)? Are 
there any limitations?
\end{verbatim}

We agree with what the reviewer remarks, and so we have created a new Section (4.4, in page XX), describing the trials that have been conducted with the first prototype of MUSES. In that section there are commented the ratio of security violations detected and how the system has influenced in the complete reduction of this number to 0.
This has been also commented in the Introduction of the paper.

The main limitation of MUSES, also commented in Section 4.4, is the need of an initial set of defined security rules to start working. The other limitation comes from the set of deployed sensors, which are the only input of data in the system, so it cannot manage security violations happened by different ways.

@@

To avoid this in part, there is a set of guidelines for the companies to build adequate security policies. For instance, the set of initially defined rules used for the trial included a blacklist of forbidden applications, a whitelist of allowed ones, and requirement of antivirus, among others.

With respect to the computational costs, in this first prototype they are around 0.3 - 0.5 seconds in the client side, if there is already a decision in the local table for the event. If there is need to ask the server to make a decision (by means of Event Correlation + Risk and Trust Analysis), the time grows up to 2 - 2.5 seconds in the worst case.
This has been also commented in Section 4.4.

% Antonio - completo con cifras que me ha dado Sergio.

### 2015-07-09 MUSES_vs_BYOD_solutions.tex
Respuestas al segundo revisor completas también
Y nuevas cosas en el artículo

>>>>

With respect to the computational costs, in this first prototype it takes less than 0.5 seconds in the client side, if there is already a decision in the local table for the happened event. Otherwise, i.e. if there is need to ask the server to make a decision (by means of Event Correlation + Risk and Trust Analysis), the time grows up to 2.5 seconds in the worst case.
